Scaffolder - Software for Reproducible Genome Scaffolding.

### 2011-03-07 src/article.tex
Spell check

The process of resolving a genome sequence and building a scaffold uses wet
laboratory methods, \emph{in silico} methods, or a combination of both. An
example of a computational method might use the existing paired-read data from
the sequencing stage where reads were produced in pairs at a known distance
apart in the genome. The occurrence of paired reads in separate contigs can be
used to probabalistically estimate the order and distance between these
contigs. Alternatively, laboratory methods may use PCR to amplify the unknown
DNA in a gap region, then use traditional Sanger sequencing to determine the
sequence of this gap. Computational methods are more preferable as they are
less costly in laboratory time and materials compared to manual gap resolution,
and use the available data that exists following sequencing
\cite{nagarajan2010}. Finally when the scaffold cannot be completely resolved,
\emph{in silico} software packages exist to suggest the likely primers
necessary for PCR amplifying the sequence in gap regions \cite{gordon2001}. \pb

@@

Repeat regions may also be responsible for multiple gaps when building a genome
sequence; tandemly repeated nucleotide regions in the genome produces multiple
reads with similar sequence. As many assembly algorithms rely on sequence
overlaps between reads to build a contig the similarity between repeat-region
reads can result in the assembly collapsing into an artifically short sequence
at this region or being ignored by more conservative assembly algorithms. Such
regions can be resolved by using algorithms to specifically reassemble the
collapsed repeat region correctly \cite{mulyukov2002,koren2010}, while
a related approach uses unassembled sequence reads matching the regions around
a scaffold gap to construct a uniquely overlapping set of reads across it
\cite{tsai2010}. \pb

@@

These described methods provide a wide array of approaches for merging contigs
into a larger, continuous scaffold sequence. Nevertheless the scaffolding
process may still require inserting additional sequences or further joining
contigs using PCR-derived sequence. This manual stage in the finishing
processmay require moving and editing large blocks of nucleotide text by hand
which is subject to human error and precludes any reproducibility of the steps
in this process. \pb

@@

Scaffolding requires joining assembled contigs and insert additional sequence
data using a combination of computational and laboratory methods. The process
of manually editing a scaffold is inherently hard to reproduce and introduces
the possibility of irreporducible techniques and/or human error. The aims of
the Scaffolder software are twofold: 1) to provide software that is easy to
install and simplifies the task of genome finishing; and 2) to facilitate
reproducibility when joining large nucleotide sequences together to create
a genome sequence. \pb

>>>>

The process of resolving a genome sequence and building a scaffold uses wet
laboratory methods, \emph{in silico} methods, or a combination of both. An
example of a computational method might use the existing paired-read data from
the sequencing stage where reads were produced in pairs at a known distance
apart in the genome. The occurrence of paired reads in separate contigs can be
used to probabilistically estimate the order and distance between these
contigs. Alternatively, laboratory methods may use PCR to amplify the unknown
DNA in a gap region, then use traditional Sanger sequencing to determine the
sequence of this gap. Computational methods are more preferable as they are
less costly in laboratory time and materials compared to manual gap resolution,
and use the available data that exists following sequencing
\cite{nagarajan2010}. Finally when the scaffold cannot be completely resolved,
\emph{in silico} software packages exist to suggest the likely primers
necessary for PCR amplifying the sequence in gap regions \cite{gordon2001}. \pb

@@

Repeat regions may also be responsible for multiple gaps when building a genome
sequence; tandemly repeated nucleotide regions in the genome produces multiple
reads with similar sequence. As many assembly algorithms rely on sequence
overlaps between reads to build a contig the similarity between repeat-region
reads can result in the assembly collapsing into an artificially short sequence
at this region or being ignored by more conservative assembly algorithms. Such
regions can be resolved by using algorithms to specifically reassemble the
collapsed repeat region correctly \cite{mulyukov2002,koren2010}, while
a related approach uses unassembled sequence reads matching the regions around
a scaffold gap to construct a uniquely overlapping set of reads across it
\cite{tsai2010}. \pb

@@

These described methods provide a wide array of approaches for merging contigs
into a larger, continuous scaffold sequence. Nevertheless the scaffolding
process may still require inserting additional sequences or further joining
contigs using PCR-derived sequence. This manual stage in the finishing
process may require moving and editing large blocks of nucleotide text by hand
which is subject to human error and precludes any reproducibility of the steps
in this process. \pb

@@

Scaffolding requires joining assembled contigs and insert additional sequence
data using a combination of computational and laboratory methods. The process
of manually editing a scaffold is inherently hard to reproduce and introduces
the possibility of irreducible techniques and/or human error. The aims of
the Scaffolder software are twofold: 1) to provide software that is easy to
install and simplifies the task of genome finishing; and 2) to facilitate
reproducibility when joining large nucleotide sequences together to create
a genome sequence. \pb
